Formalizing Stack Safety as a Security Property

### 2021-07-08 submission3/main.tex
Response to comments

The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often involve trading off protection and performance.\bcp{I'm a
  little uncomfortable about that point: to the extent that applying these
  mechanisms involves trading off performance and protection, our properties
will not apply at all!}
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, as a basis both for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

@@

\bcp{Have we made the diagrams comprehensible when printed in
  B/W or viewed by a colorblind person?  If so we should say that the colors
are just extra hints.}%
Colored boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of intended access
rights.\bcp{should we say here that it is the job of an enforcement mechanism to
  manage and enforce these access permissions?} Such a domain
can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
figures\leo{No outside?}. For our purposes, \(\outside\) will apply to global
variables and code. Domains are part of the {\em context} of a machine state.
(We will introduce some other parts of the context as needed.)

@@

\section{Setting}
\label{sec:setup}

\bcp{I am having trouble deciding what this section should be called.
  Probably this means I don't really understand its purpose...}
\sna{Propose merging it with next section like so.}
\leo{Machine, Enforcement, and Threat Model?}

\bcp{And: Needs an intro sentence.}

@@

Some calling convention choices, such as callee-save registers and argument
passing on the stack, require natural enhancements to our base properties. We
present such enhancements in Section~\ref{sec:ext}, after we formalize the
essence of stack safety. Other plausible enhancements
include:

@@

Runtime code generation would be much more challenging to support, as
it requires dynamic code annotations. Parallelism would require
fundamental changes to the system model. So, we consider these features
fully out of scope.\leo{Shouldn't coroutines be mentioned here?}

@@

\paragraph*{Policies}\leo{Change this to ``enforcement model'' if we go with that title}
We model the base machine's behavior as a step function over machine states,
and a machine
enhanced with an enforcement mechanism, or {\em policy} for short, as a
partial step function over machine states extended with some
auxiliary state, refining the base machine. When the policy-enhanced machine
does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
action that would violate the security property;
our properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

@@

The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
gives the trace of observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
initial observation, a machine-trace, and an observation trace:\bcp{This
  would read better and take less space if all three rules were squished
  onto one line.}

\begin{minipage}[b]{.6\textwidth}
  \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
                {\(\mpcstatename' \hookrightarrow \MPCT\)}
                {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}%
\end{minipage}
\begin{minipage}[b]{.35\textwidth}
  \judgment{}
           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
\end{minipage}

\begin{minipage}{.35\textwidth}
  Then we start with a \(\tau\) observation:
\end{minipage}
\begin{minipage}{.6\textwidth}
\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
         {\(\obsof(\MPCT) = \obsT\)}
\end{minipage}

@@

\definition
For each of the stepwise, deferred, and observational versions of
stack integrity in the simple model, the equivalent property in the
passing model\bcp{Is something missing here?}
%
{\bf Observational stack integrity} with callee-saves registers substitutes the above
\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
%
{\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).

>>>>

The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate, and the behaviors they prevent
can vary. Perhaps one conservatively overapproximates the ideal stack safety,
while another fails to measure up. In order to say for certain,
we need a precise, generic, and formal specification for stack
safety, as a basis both for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

@@

Labeled boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of intended access
rights. Unless otherwise marked, sequential addresses retain the previous domain,
as hinted by their colors. Domains are descriptive; it is up to the enforcement
mechanism to ensure that the system adheres to these permissions. Such a domain
can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
figures. Not shown are global variables and code, which would be labeled \(\outside\).
Domains are part of the {\em context} of a machine state.
(We will introduce some other parts of the context as needed.)

@@

\section{Machines, Enforcement, and Threat Model}
\label{sec:setup}

Before introducing the model itself, we must situate it in terms of our
assumptions about machines, enforcement mechanisms, and threats.

@@

Some calling convention choices, such as callee-save registers and argument
passing on the stack, require natural enhancements to our base properties. We
present these enhancements, as well as a simple coroutine model,
in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
Other plausible enhancements include:

@@

Runtime code generation would be much more challenging to support, as
it requires dynamic code annotations. Parallelism beyond the current
coroutine model would require much more substantial changes to the machine
model. So, we consider these features out of scope.

@@

\paragraph*{Enforcement Model}
We model the base machine's behavior as a step function over machine states,
and a machine
enhanced with an enforcement mechanism, or {\em policy} for short, as a
partial step function over machine states extended with some
auxiliary state, refining the base machine. When the policy-enhanced machine
does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
action that would violate the security property;
our properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

@@

The ``observations-of'' operator, written \(\obsof\), takes a machine trace and
gives the trace of observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsfor}\) that relates an
initial observation, a machine-trace, and an observation trace; the helper is
then invoked with an initial \(\tau\).

\begin{minipage}[b]{.55\textwidth}
  \judgmentthree{\(\mpcstatename \stepstocon[\obs'] \mpcstatename'\)}
                {\(\mpcstatename' \hookrightarrow \MPCT\)}
                {\(\mathit{obsfor} ~ \obs' ~ \MPCT = \obsT\)}
                {\(\mathit{obsfor} ~ \obs ~ \notfinished{\mpcstatename}{\MPCT} = \notfinished{\obs}{\obsT}\)}%
\end{minipage}
\begin{minipage}[b]{.22\textwidth}
  \judgment{}
           {\(\mathit{obsfor} ~ \obs ~ \mpcstatename = \obs\)}
\end{minipage}
\begin{minipage}[b]{.22\textwidth}
\judgment{\(\mathit{obsfor} ~ \tau ~ \MPCT = \obsT\)}
         {\(\obsof(\MPCT) = \obsT\)}
\end{minipage}

@@

\definition
%
{\bf Observational stack integrity} with callee-saves registers substitutes the above
\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
%
{\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).
